Supplement for: Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming

### 2014-11-11 manuscripts/manuscript.Rmd
supplement must be built in same command after manuscript.

We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to
infer posterior distributions of the parameters of the GP (Figure
S3, code in appendix). Since the posterior distributions differ
substantially from the priors (Figure S3), most of the information in the posterior comes from the data rather than the
prior belief.

>>>>

We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to
infer posterior distributions of the parameters of the GP (Figure
S4, code in appendix). Since the posterior distributions differ
substantially from the priors (Figure S4), most of the information in the posterior comes from the data rather than the
prior belief.

### 2014-11-11 manuscripts/supplement.Rmd
supplement must be built in same command after manuscript.

```{r knit, include=FALSE}
library("methods")
library("rmarkdown")
render("manuscript.Rmd")
```

@@

@@

# Training data #

@@

# Model performance outside the predicted range

@@

```{r figure_S1, fig.width=8, fig.height=6, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP. Points show the stock size simulated by the true model.  Overlay shows the range of states predicted by each model, based on the state observed in the previous time step. The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).  Note that because it does not assume a parametric form but instead relies more directly on the data, the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
y <- numeric(8)
y[1] <- 4.5
for(t in 1:(length(y)-1))
      y[t+1] = z_g() * f(y[t], h=0, p=p)
# predicts means, does not reflect uncertainty estimate!
crash_data <- step_ahead_posteriors(y)
crash_data <- subset(crash_data, variable %in% c("GP", "Allen", "Ricker", "Myers"))
ggplot(crash_data) + 
  geom_boxplot(aes(as.factor(as.integer(time)), value, 
                   fill = variable, col=variable), 
               outlier.size=1, position="identity") + 
#  geom_line(aes(time, value, col = variable, 
#            group=interaction(L1,variable))) + 
  geom_point(aes(time, stock), size = 3) + 
  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
  facet_wrap(~variable) + 
  theme(legend.position="none") + xlab("time") + ylab("stock size")

@@

```{r figure_S2, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
gp_assessment_plots$traces_plot
```

```{r figure_S3, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
gp_assessment_plots$posteriors_plot
write.csv(gp_dat,"components/data/figureS2-3.csv")
```

```{r figure_S4, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
ggplot(ricker_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S5, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(ricker_posteriors, "components/data/figureS4-5.csv")
```

```{r figure_S6, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
ggplot(myers_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S7, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
ggplot(myers_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(myers_posteriors, "components/data/figureS6-7.csv")
```

```{r figure_S8, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
ggplot(allen_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S9, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
ggplot(allen_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(allen_posteriors, "components/data/figureS8-9.csv")
```

# Further sensitivity analysis #

@@

```{r figure_S10, fig.height=6, fig.width=10, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.  Histograms shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters. Values of 1 indicate optimal performance. Columns indicate different models, rows different noise levels, and colors indicate the parameter set used. Grouped over stochastic replicates applying the contol policy and stochastic replicates of training data generated from the model indicated, see raw data for details. Randomly chosen parameter values for the models shown in tables below."}
ggplot(sensitivity_dat) + 
  geom_histogram(aes(value, fill=parameters)) + 
  xlim(0,1.0) + 
  theme_bw() + 
  xlab("value as fraction of the optimal") + 
  facet_grid(noise~model)
write.csv(sensitivity_dat, "components/data/figureS10.csv") 
```

```{r AllenSetsTable, results="asis"}
pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S9." )
```
```{r MyersSetsTable, results="asis"}
pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S9." )
```

>>>>

```{r knit, include=FALSE, eval=FALSE}
library("methods")
library("rmarkdown")
render("manuscript.Rmd")
```

@@

## Dryad Data Archive ##

@@

## Training data description ##

@@

# Model performance outside the predicted range (Fig S1) #

@@

```{r figure_S1, fig.width=6, fig.height=4, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP. Points show the stock size simulated by the true model.  Overlay shows the range of states predicted by each model, based on the state observed in the previous time step. The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).  Note that because it does not assume a parametric form but instead relies more directly on the data, the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
y <- numeric(8)
y[1] <- 4.5
for(t in 1:(length(y)-1))
      y[t+1] = z_g() * f(y[t], h=0, p=p)
# predicts means, does not reflect uncertainty estimate!
crash_data <- step_ahead_posteriors(y)
crash_data <- subset(crash_data, variable %in% c("GP", "Allen", "Ricker", "Myers"))
ggplot(crash_data) + 
  geom_boxplot(aes(as.factor(as.integer(time)), value, 
                   fill = variable, col=variable), 
               outlier.size=1, position="identity") + 
#  geom_line(aes(time, value, col = variable, 
#            group=interaction(L1,variable))) + 
  geom_point(aes(time, stock), size = 3) + 
  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
  facet_wrap(~variable) + 
  theme(legend.position="none") + xlab("time") + ylab("stock size")

@@

# Further sensitivity analysis (Fig S2) #

@@

```{r figure_S2, fig.height=6, fig.width=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.  Histograms shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters. Values of 1 indicate optimal performance. Columns indicate different models, rows different noise levels, and colors indicate the parameter set used. Grouped over stochastic replicates applying the contol policy and stochastic replicates of training data generated from the model indicated, see raw data for details. Randomly chosen parameter values for the models shown in tables below."}
ggplot(sensitivity_dat) + 
  geom_histogram(aes(value, fill=parameters)) + 
  xlim(0,1.0) + 
  theme_bw() + 
  xlab("value as fraction of the optimal") + 
  facet_grid(noise~model)
write.csv(sensitivity_dat, "components/data/figureS2.csv") 
```

```{r AllenSetsTable, results="asis"}
pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S2." )
```
```{r MyersSetsTable, results="asis"}
pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S2." )
```

## GP MCMC (Fig S3-4) ##

```{r figure_S3, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options", fig.width=4, fig.height=4}
gp_assessment_plots$traces_plot
```

```{r figure_S4, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red.", fig.width=4, fig.height=4}
gp_assessment_plots$posteriors_plot
write.csv(gp_dat,"components/data/figureS3-4.csv")
```

\newpage
\newpage

## Ricker Model MCMC (Fig S5-6) ##

```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
ggplot(ricker_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S6, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=4, fig.height=4}
ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(ricker_posteriors, "components/data/figureS5-6.csv")
```

\newpage
\newpage

## Myers Model MCMC (Fig S7-8) ##

```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
ggplot(myers_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S8, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=4, fig.height=6}
ggplot(myers_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(myers_posteriors, "components/data/figureS7-8.csv")
```

\newpage 
\newpage

## Allen Model MCMC (Fig S9-10) ##

```{r figure_S9, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
ggplot(allen_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free_y", ncol=1)
```

```{r figure_S10, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=4, fig.height=6}
ggplot(allen_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(allen_posteriors, "components/data/figureS9-10.csv")
```
